---
title: "2025-01-07 R ex"
---

## The Course Script

JF provided two function: `makeweb()` to simulate a web dataset using `rnorm()` and `sampleweb()` function to generate resampling from generated web. Original R script is stored under "/Resources"

```{r}
# small simulation for AREES paper, abun&network structure, chicken-or-egg

# ---- functions etc. ----
library(bipartite)

# load two functions from Fründ et al. 2016 Oikos
makeweb <- function(specpar = 1, Nbee=4, Nplant=4, specvar=0, planttrait_n="single"){    
  # function to generate web with defined "specialization"; now the generalized version with multiple methods
  # specpar is the specialization parameter (always increases with specialization), specvar is within network variation of specialization
  # specpar is the spec. parameter; exponent in "power", quad.slope in "parabolic", 1/sd in "normal" 
  # now simplified to method "traitmatch" with nicheshape "normal"!!
  # planttrait_n is used for method "traitmatch": "single" is one value for each plant per web, "multi" is one value for each plant per bee 
  # if (specvar != 0) warning('specvar currently ignored')  # currently, all bees in web have same expected specialization
  #! specvar now implemented: sd-factor; specpar is a random variate of specpar
  fun_pref <- function(traitdif){
    specpar.new <- rnorm(1, specpar, specpar*specvar)
    if (specpar.new <= 0) specpar.new <- 0.1
    prefs <- dnorm(traitdif, mean=0, sd=1/specpar.new)
    prefs
  }
  if (planttrait_n == "single"){
    planttrait <- runif(Nplant)    
    web <- replicate(Nbee, fun_pref(abs(runif(1) - planttrait)))
  }
  if (planttrait_n == "multi"){
    web <- replicate(Nbee, fun_pref(abs(runif(1) - runif(Nplant))))
  }      
  web <- web / matrix(colSums(web),nrow=Nplant,ncol=Nbee,byrow=TRUE)    # standardize link weights to probability;
  web
}

sampleweb <- function(web,obsperbee,method='perbee'){   # subsampling a network; either fixed Nobs per bee or fixed Nobs per web
  if(method=='perbee') {
    sampledweb <- sapply(1:ncol(web), function(j)  {
      web.j <- web[,j]
      table(sample(factor(1:length(web.j)),obsperbee,prob=web.j,replace=TRUE))
    })
  }
  if(method=='perweb'){
    Nobs <- obsperbee * ncol(web)
    sampledweb.vect <- sample(as.factor(1:length(web)), size=Nobs, prob=as.numeric(web), replace=TRUE)
    sampledweb <- matrix(as.numeric(table(sampledweb.vect)), nrow=nrow(web))       # puts back into matrix format
  }
  dimnames(sampledweb) <- dimnames(web)
  sampledweb
}
```

### Exercise #1 How do sampling effort effects network structure / patterns?

- Use simulation functions to establish the effect of sampling intensity on a network index of your choice
- `plot(indexvalue ~ Nobs)` # Nobs = observation * ncol(web)
- use `sampleweb(method = "perweb")`

Create a "parent" web dataset from which the resampling will be drawed.

```{r}
parent.web <- makeweb(Nbee = 10, Nplant = 10, specpar = 3)
parent.web
```

A single subsampling from the parent web data with `obsperbee = 8`.

```{r}
sample.web <- sampleweb(parent.web, obsperbee = 40, method = "perweb")
sample.web
```

Calculate a network index with the `sample.web`

```{r}
networklevel(sample.web, index = "NODF")
```

To see the response of the index to sample size of the web, we repeat this process with different web size by changing the `obsperbee` in `sample.web()`.

```{r}
library(tidyverse)
indices <- c("NODF", "connectance")
indices.p.size <- map(indices, \(y) {
    print(y)
    index.p.size <- map_dfr(c(1:100), \(x) {
        sample.web <- sampleweb(parent.web, obsperbee = x, method = "perweb")
        web.size <- sum(sample.web)
        index <- networklevel(sample.web, index = y)
        return(list(Size = web.size, Index = index, name = y))
    })
}) %>% bind_rows()


# index.p.size
indices.p.size <- indices.p.size %>%
    mutate(name = factor(name))
plot <- indices.p.size %>% ggplot() +
    geom_point(aes(x = Size, y = Index, color = name))
plot
```

### Exercise #2 Fancy Plots using Bluewiese data

A WIP [vignette](/Resources/How_to_plot.html) and functionality of future `bipartite::` package with fancy plotting functionality. 

Repeating same import and basic data cleaning as yesterday:

```{r}
Wiese <- read.csv("Data/Wiese.csv")
Wiese <- Wiese[Wiese$`Bl.te` != "keine",]
weblist <- frame2webs(Wiese, varnames = c("Bl.te", "Beobachtetes.Insekt", "Plot"))
str(weblist)
```

`weblist` contains networks of 3 plots. A recap with old `plotweb()` function on those:

```{r}
walk(weblist, \(x) plotweb(x))
```

Use new version `plotweb_v2()` to plot BP1 network:

```{r}
plotweb_v2(weblist[[1]], mar = c(0, 0, 0, 0), text_size = 1, horizontal = T)
```

```{r}
plot2webs_v2(weblist[[1]], weblist[[2]], text_size = 0.5, horizontal = T)
```